Hydras \& Co.

### 2022-03-12 doc/hydras.tex
re-reading pages 9-21

\newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
\newtheorem{gaiahydra}{\gaiasign}

@@

\begin{itemize}
\item Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
Although the mathematical contents of this 
paper are quite advanced, the rules of hydra battles are very easy to understand. There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with. See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.

Let us underline the analogy between hydra battles and interactive theorem proving. Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.

@@

\begin{scope}[xshift=-6cm]
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=2.2cm] (pock) { Pocklington };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=2.5cm] (additions) { Additions };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=2cm] (cantor) { Cantor };

\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=3.2cm] (cats) { Categories in ZFC };
\end{scope}

\begin{scope}[yshift=-1cm,xshift=0.25cm]
  \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=1.8cm] (mathcomp) { MathComp };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=1.5cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
\end{scope}

@@

Quotations of \coq{} source  from Standard Library and code which lead to compilation error or too long computations are displayed as follows.

@@

>>>>

\newcommand{\gaiacompat}{\textcolor{blue}{Gaia}}
\newcommand{\gaiasign}{\texorpdfstring{\mycircled[cyan!40]{{\mbox{\color{blue}{\scriptsize{G}}}}}}{Gaia-Hydra}}
\newtheorem{gaiahydra}{\gaiasign}

@@

\begin{itemize}
\item Hydra games (a.k.a. \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians:
L. Kirby and J. Paris~\cite{KP82}: \emph{Accessible Independence Results for Peano Arithmetic}. 
Although the mathematical contents of this 
paper are quite advanced, the rules of hydra battles are very easy to understand\footnote{Let us underline the analogy between hydra battles and interactive theorem proving. Hercules is the user (you!), and hydra's heads are the subgoals: you may think that applying a tactic would solve a subgoal, but it results often in the multiplication of such tasks.}.
There are now several sites on the Internet where you can find tutorials on hydra games, together with simulators you can play with. See, for instance, the blogpost and source code written by Andrej Bauer~\cite{bauer2008,BauerHydra}.

@@

\begin{scope}[xshift=-6.5cm]
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm] (goedel) { Goedel };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,left of=goedel, node distance=1.9cm] (pock) { Pocklington };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=goedel, node distance=1.8cm] (additions) { Additions };
\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=additions, node distance=1.8cm] (cantor) { Cantor };

\node[rectangle, dotted,draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=cantor, node distance=2.2cm] (cats) { Categories in ZFC };
\end{scope}

\begin{scope}[yshift=-1cm,xshift=-4.25cm]
  \node[rectangle, draw=black,fill=blue!20,minimum height=0.5cm,minimum width=1.5cm] (paramcoq) { Paramcoq };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,right of=paramcoq, node distance=5.8cm] (mathcomp) { MathComp };
\node[rectangle, draw=black,minimum height=0.5cm,minimum width=1.5cm,node distance=1.5cm,left of=paramcoq, node distance=4.5cm] (equations) { Equations };
\end{scope}

@@

Quotations of \coq{} source  from others libraries (\coq's standard library, borrowed plug-ins)  are displayed as follows.

@@

We use also verbatim code inclusions when the examples lead to very long computations.

### 2022-03-12 doc/part-hydras.tex
re-reading pages 9-21

\subsection*{Compatibility with \gaia\cite{Gaia} (in progress)}
The \gaia project~\cite{Gaia} by Jos\'e Grimm aimed to formalize mathematics in \coq  in
the style of Nicolas Bourbaki. It contains many definitions and results about ordinal numbers. In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa. Remarks about compatibility with \gaia signalled with the picture \gaiasign.

@@

\begin{itemize}
\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
\item Extension to $\Gamma_0$ (in Veblen normal form)
\item More lemmas about hierarchies of rapidly growing
  functions, and their relationship 
    with primitive recursive functions and provability in Peano arithmetic 
(following~\cite{KS81, KP82}).
\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
\end{itemize}

>>>>

\subsection*{Compatibility with \gaia (in progress)}
The \gaia project~\cite{Gaia} by Jos\'e Grimm, Alban Quadrat,  and Carlos Simpson,
aims  to formalize mathematics in \coq  in
the style of Nicolas Bourbaki. It contains many definitions and results about ordinal numbers. In Chapter~\ref{gaia-chapter}, we present some modules which allow \HydrasLib' users to apply lemmas proven in \gaia, and vice versa. Remarks about compatibility with \gaia are signalled with the picture \gaiasign.

@@

\begin{itemize}
\item Semiautomatic tactics for proving inequalities $\alpha < \beta$, even when $\alpha$ and $\beta$ are not closed terms.
%\item Extension to $\Gamma_0$ (in Veblen normal form)
\item More lemmas about hierarchies of rapidly growing
  functions, and their relationship 
    with primitive recursive functions and provability in Peano arithmetic 
(following~\cite{KS81, KP82}).
\item From \coq's point of view, this development could be used as an illustration of the evolution of the software, every time new libraries and sets of tactics could help to simplify the proofs.
\end{itemize}
